Hierarchical storage management in user space for neuroimaging applications

### 2024-03-13 paper/sea-neuro/paper-neuro.tex
Removed more duplicates

For both the pipelines and the datasets, we noticed that obtained speedups
decreased with increased parallelism. We did not modify the default parallelism
level of the applications in our experiments. As a result, each application
process was likely attempting to utilize all cores, resulting in longer wait times
with the increase in parallelism. With this unintentional additional wait time that
process was likely attempting to utilize all cores, resulting in longer wait times
with the increase in parallelism. With this unintentional additional wait time that
is added to the applications due to compute contention, the benefits of using
Sea becomes more limited.

@@

  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
  shared with other users,
  using the same datasets and pipelines as for the controlled cluster. 
  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
  shared with other users,
  using the same datasets and pipelines as for the controlled cluster. 
  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
  show that although performance in production clusters was highly variable,
  we could still obtain important speedups by using Sea. It is very likely that when we
  executed our experiments without flushing
  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
  resulting in Sea and Baseline performing
  quite similarly to each other.

  Our results with flushing enabled confirmed this theory. When running with
  flushing enabled, which had the additional overhead of having to
  ensure that all the data was copied to Lustre, we noticed that occasionally, we
  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
  Our results with flushing enabled confirmed this theory. When running with
  flushing enabled, which had the additional overhead of having to
  ensure that all the data was copied to Lustre, we noticed that occasionally, we
  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
  performance was degraded. The
  maximum speedup observed when running on the production cluster with flushing enabled was
  11$\times$ with the AFNI pipeline processing 1 HCP image. The maximum speedup for the SPM pipeline was
  3.2$\times$, processing 8 ds001545 images.

@@

    Although Sea usage did result in performance degradation at times, the magnitude
    of the degradation was less than that of possible speedups
    obtained. Our controlled experiments demonstrated that using Sea with a
    of the degradation was less than that of possible speedups
    obtained. Our controlled experiments demonstrated that using Sea with a
    degraded Lustre file system can result in average speedups of up to
    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
    less important that the potential speedups that can arise.

@@

    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
     when the data is written at a rate which exceeds the rate at which the
    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
     when the data is written at a rate which exceeds the rate at which the
    compute node's page cache can flush results to Lustre. In our experiments,
    the pipeline which consisted of both the shortest duration and largest
    output size was 16 AFNI pipeline processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
    approximately 100~GB of data, the controlled cluster would have had
    approximately 44~GB of dirty cache available per Lustre Object Storage Target (OST) and a maximum amount of
    100~GB of page cache. Since all writes would not have occurred at the same
    time and computation time was extended as a result of CPU contention, even
    our controlled experiments did not showcase the benefits of Sea in a highly
    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
    the only way to augment data intensiveness of the applications is
    our controlled experiments did not showcase the benefits of Sea in a highly
    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
    the only way to augment data intensiveness of the applications is
    through increased parallelism which is limited by the number of available
    cores. As a result, we could only demonstrate the speedups brought by Sea when
    Lustre performance had been degraded.
    cores. As a result, we could only demonstrate the speedups brought by Sea when
    Lustre performance had been degraded.

    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
    In such cases, the performance bottleneck is the
    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
    In such cases, the performance bottleneck is the
    CPU time, which Sea does not address. Moreover, depending on the
    application's I/O patterns, it is possible that there will be sufficient
    page cache space and time to flush all the results asynchronously
    to Lustre. However, even with minimal speedups, Sea can still be useful in
    these cases to limit the number of new files created on Lustre, which can be
    particularly useful in instances where users have strict file limits.

@@

    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
    where we had
    control over the global cluster usage and were able to demonstrate that
    significant speedups could be obtained when all the Lustre storage devices were
    busy writing other data concurrently. However, even these experiments were
    limited in demonstrating potential speedups with Sea as there were many
    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
    where we had
    control over the global cluster usage and were able to demonstrate that
    significant speedups could be obtained when all the Lustre storage devices were
    busy writing other data concurrently. However, even these experiments were
    limited in demonstrating potential speedups with Sea as there were many
    controlled variables, some of which may vary from cluster to cluster. For
    example, the production cluster had a total of 977 nodes and only 38 disks associated with
    their Lustre scratch file system. We can imagine that if 900 of the nodes
    are busy writing to the scratch file system, we would observe a speedup
    larger than what has been reported on our dedicated cluster with only 8 compute
    nodes and 44 disks associated to Lustre. However, these are not the only
    factors that can affect reported results, others can include when the
    experiment were executed, what kind of applications were being executed
    alongside the experiment, scratch file system usage during experiments, and
    what were the application read and write patterns.

@@

    On the dedicated cluster, we ran the experiments with and without
    Lustre degradation, which was induced by ``busy writers''. Without busy
    writers, the fMRI preprocessing pipelines had exclusive access to the
    cluster. With busy writers, the fMRI preprocessing pipelines were executed
    alongside an Apache Spark application that continuously read and wrote approximately
    writers, the fMRI preprocessing pipelines had exclusive access to the
    cluster. With busy writers, the fMRI preprocessing pipelines were executed
    alongside an Apache Spark application that continuously read and wrote approximately
    1000$\times$ \SI{617}{\mebi\byte} blocks using 64 threads, with a 5 seconds sleep between reads
    and writes. In our experiments, we either had 6 nodes each executing the
    Spark application or no busy writers.

    Our experiments generally did not use prefetching, flushing or eviction. The
    only pipeline configured to prefetch was SPM, as the input data was otherwise
    read through a memmap (i.e. only loading necessary portions of the file to memory) 
    with Lustre. Flushing would normally be necessary with
    only pipeline configured to prefetch was SPM, as the input data was otherwise
    read through a memmap (i.e. only loading necessary portions of the file to memory) 
    with Lustre. Flushing would normally be necessary with
    preprocessing as the user would always require the output data. To
    investigate the impacts of flushing on these pipelines, we performed a
    separate set of experiments where AFNI and SPM would flush all data
    produced, on the production cluster.

@@

    All experiments were executed using 1, 8 and 16 processes.
    Each process consisted of a single application call processing a single fMRI image.
    All experiments were executed using 1, 8 and 16 processes.
    Each process consisted of a single application call processing a single fMRI image.

>>>>

For both the pipelines and the datasets, we noticed that obtained speedups
decreased with increased parallelism. We did not modify the default parallelism
level of the applications in our experiments. As a result, each application
process was likely attempting to utilize all cores, resulting in longer wait times
with the increase in parallelism. With this unintentional additional wait time that
is added to the applications due to compute contention, the benefits of using
Sea becomes more limited.

@@

  We also benchmarked Sea on a production cluster of the Digital Alliance of Canada
  shared with other users,
  using the same datasets and pipelines as for the controlled cluster. 
  Figures~\ref{fig:seaneuro:beluga-noflush} and \ref{fig:seaneuro:beluga-wflush}
  show that although performance in production clusters was highly variable,
  we could still obtain important speedups by using Sea. It is very likely that when we
  executed our experiments without flushing
  (Figure~\ref{fig:seaneuro:beluga-noflush}), Lustre performance was not degraded, 
  resulting in Sea and Baseline performing
  quite similarly to each other.

  Our results with flushing enabled confirmed this theory. When running with
  flushing enabled, which had the additional overhead of having to
  ensure that all the data was copied to Lustre, we noticed that occasionally, we
  obtained very large speedups with both the SPM and AFNI pipelines, suggesting that Lustre
  performance was degraded. The
  maximum speedup observed when running on the production cluster with flushing enabled was
  11$\times$ with the AFNI pipeline processing 1 HCP image. The maximum speedup for the SPM pipeline was
  3.2$\times$, processing 8 ds001545 images.

@@

    Although Sea usage did result in performance degradation at times, the magnitude
    of the degradation was less than that of possible speedups
    obtained. Our controlled experiments demonstrated that using Sea with a
    degraded Lustre file system can result in average speedups of up to
    2.5$\times$. Such results indicate that the risk of a slowdown with Sea are
    less important that the potential speedups that can arise.

@@

    Sea's benefit is maximal when it is executed on data-intensive pipelines, that is,
     when the data is written at a rate which exceeds the rate at which the
    compute node's page cache can flush results to Lustre. In our experiments,
    the pipeline which consisted of both the shortest duration and largest
    output size was 16 AFNI pipeline processes preprocessing 16 HCP fMRI images. While the concurrent processes wrote, in total,
    approximately 100~GB of data, the controlled cluster would have had
    approximately 44~GB of dirty cache available per Lustre Object Storage Target (OST) and a maximum amount of
    100~GB of page cache. Since all writes would not have occurred at the same
    time and computation time was extended as a result of CPU contention, even
    our controlled experiments did not showcase the benefits of Sea in a highly
    data-intensive scenario. Since HCP is one of the largest fMRI datasets,
    the only way to augment data intensiveness of the applications is
    through increased parallelism which is limited by the number of available
    cores. As a result, we could only demonstrate the speedups brought by Sea when
    Lustre performance had been degraded.

    Sea cannot speedup applications that are very compute-intensive, as is the case with the FSL Feat pipeline.
    In such cases, the performance bottleneck is the
    CPU time, which Sea does not address. Moreover, depending on the
    application's I/O patterns, it is possible that there will be sufficient
    page cache space and time to flush all the results asynchronously
    to Lustre. However, even with minimal speedups, Sea can still be useful in
    these cases to limit the number of new files created on Lustre, which can be
    particularly useful in instances where users have strict file limits.

@@

    Our results expose the complexities of performance predictions on HPC clusters. We presented results on a dedicated cluster
    where we had
    control over the global cluster usage and were able to demonstrate that
    significant speedups could be obtained when all the Lustre storage devices were
    busy writing other data concurrently. However, even these experiments were
    limited in demonstrating potential speedups with Sea as there were many
    controlled variables, some of which may vary from cluster to cluster. For
    example, the production cluster had a total of 977 nodes and only 38 disks associated with
    their Lustre scratch file system. We can imagine that if 900 of the nodes
    are busy writing to the scratch file system, we would observe a speedup
    larger than what has been reported on our dedicated cluster with only 8 compute
    nodes and 44 disks associated to Lustre. However, these are not the only
    factors that can affect reported results, others can include when the
    experiment were executed, what kind of applications were being executed
    alongside the experiment, scratch file system usage during experiments, and
    what were the application read and write patterns.

@@

    On the dedicated cluster, we ran the experiments with and without
    Lustre degradation, which was induced by ``busy writers''. Without busy
    writers, the fMRI preprocessing pipelines had exclusive access to the
    cluster. With busy writers, the fMRI preprocessing pipelines were executed
    alongside an Apache Spark application that continuously read and wrote approximately
    1000$\times$ \SI{617}{\mebi\byte} blocks using 64 threads, with a 5 seconds sleep between reads
    and writes. In our experiments, we either had 6 nodes each executing the
    Spark application or no busy writers.

    Our experiments generally did not use prefetching, flushing or eviction. The
    only pipeline configured to prefetch was SPM, as the input data was otherwise
    read through a memmap (i.e. only loading necessary portions of the file to memory) 
    with Lustre. Flushing would normally be necessary with
    preprocessing as the user would always require the output data. To
    investigate the impacts of flushing on these pipelines, we performed a
    separate set of experiments where AFNI and SPM would flush all data
    produced, on the production cluster.

@@

    All experiments were executed using 1, 8 and 16 processes.
    Each process consisted of a single application call processing a single fMRI image.
